Convergence rate of MAP estimates \\ for the exponential family

### 2021-10-11 aistats.tex
remove fancy maths commands
They made my local compilation fail because I am using pdflatex 2015

% my packages
\usepackage{math_commands}
% some custom math commands
\newtheorem{proposition}{Proposition}
\newcommand*{\expect}[2][]{\ensuremath{\mathbb{E}_{#1} \left[ #2 \right] }} % expectation operator

@@

\subsection{Locally Quadratic Case}
All Bregman divergences are locally quadratic. 
Under some assumptions, we can quantify  when this quadratic behavior kicks in.
For instance {\bf when $\conj$ is self-concordant} \citep[Ch.4.1]{nesterov2003introductory}. 
This is true when $T$ is 1-dimensional and $\logpart$ is self-concordant 
-- e.g. exponential distribution, 
Gaussian\fdk{with known variance?} and Laplace distributions with known mean,
Pareto distribution with known minimum value, 
Weibull distribution with known shape $k$
-- or when $T$ lives in a compact \citep{bubeck2015entropic}. Then the bound
\begin{align}
	 \bregmanconj(\m^*,\m) \leq \norm{\m^*-\m}_{\mF^*}^2
\end{align}
holds whenever $\norm{\m^*-\m}_{\mF^*} < 0.21$.\footnote{The value of $x$ such that $x^2 \geq -\frac{x}{1-x} - \log(1 - \frac{x}{1-x})$}
Similarly, \citet{kakade2010learning} assumes a bound on all higher order moments at $\nat^*$ to quantify when $\logpart$ starts behaving quadratically, but their rate does not directly apply to $\conj$ and $\m$.

@@

\subsection{Bias-Variance Decomposition}
In both examples (add link), the upper bound takes the form $O(\inv{n}) + O(\frac{\text{bias}}{n^2})$, reminding us of the bias-variance decomposition. In fact, it is possible to write such a decomposition for any Bregman divergence
Let $\tilde \theta_n := \expect{\hat \theta_n}$ be the expectation of the MAP in primal space, and $\tilde \m_n = \nabla \logpart(\tilde \theta_n )$ be the corresponding mean parameter.
As described by \citet[Theorem 0.1]{pfau2013generalized}, the  expected Bregman decomposes like
\begin{align}
	\expect{\bregmanconj(\m^* ; \hat \m_n)} 
	&= {\bregmanconj(\m^* ; \tilde \m_n)}
	+ {\expect{\bregmanconj(\tilde \m_n ; \MAPm)}}
	\label{eq:primal_pivot}
\end{align}

@@

Relative smoothness has renewed interest for analyses of mirror descent
without relying on strong-convexity in an underlying norm.
Recent works have made attempts to generalize the assumptions 
that leads to the convergence of mirror descent 
with strongly-convex reference functions, 
such as a uniform bound on the variance of the stochastic gradients $\tilde g(\theta)$,
$\expect{\norm{\nabla f(\theta) - \tilde g(\theta)}^2} \leq \sigma^2$, 
or only requiring this bounded variance at the minimum $\theta_*$.
However, they do not yet address the problem satisfactorily.
We present those generalization here 
and they are summarized in the context of maximum likelihood estimation 
in \cref{tbl:assumptions}.

\subsection{Strongly-convex reference function}

@@

{\bf For a maximum likelihood problem, }
where $\tilde g(\theta) = \m - x$ 
where $x \sim p(x\cond\theta_*)$,
and $\nabla f(\theta) = \m - \m_*$,
\aligns{
	&
	\mathbb{E}_{x}\bigg[
	\cB_{A^*}\paren{(1-\eta) \m + \eta x, (1-\eta)\m + \eta \m_*}
	\\
	&\hphantom{\mathbb{E}_{\tilde g_t}[}
	\cB_{A^*}\paren{(1-\eta)\m + \eta \m_*, (1-\eta) \m + \eta x}
	\bigg]
	\leq 
	\eta C.
}
\fdk{Could work with decreasing step-sizes}

\fdk{How do we show that this fails?}

>>>>

% my packages
\usepackage{math_commands}
% some custom math commands
\newtheorem{proposition}{Proposition}
\newcommand*{\expect}[2][]{\ensuremath{\mathbb{E}_{#1} \left[ #2 \right] }} % expectation operator
\newcommand*{\expecta}[2][]{\ensuremath{\mathbb{E}_{#1} [ #2 ] }} % expectation operator

@@

\subsection{Locally Quadratic Case}
All Bregman divergences are locally quadratic. 
Under some assumptions, we can quantify  when this quadratic behavior kicks in.
For instance {\bf when $\conj$ is self-concordant} \citep[Ch.4.1]{nesterov2003introductory}. 
This is true when $T$ is 1-dimensional and $\logpart$ is self-concordant 
-- e.g. exponential distribution, 
Gaussian\fdk{with known variance?} and Laplace distributions with known mean,
Pareto distribution with known minimum value, 
Weibull distribution with known shape $k$
-- or when $T$ lives in a compact \citep{bubeck2015entropic}. Then the bound
\begin{align}
	 \bregmanconj(\m^*,\m) \leq \norm{\m^*-\m}_{\mF^*}^2
\end{align}
holds whenever $\norm{\m^*-\m}_{\mF^*} < 0.21$.\fdk{where does this $0.21$ come from?}
Similarly, \citet{kakade2010learning} assumes a bound on all higher order moments at $\nat^*$ to quantify when $\logpart$ starts behaving quadratically, but their rate does not directly apply to $\conj$ and $\m$.

@@

\subsection{Bias-Variance Decomposition}
In both examples (add link), the upper bound takes the form $O(\inv{n}) + O(\frac{\text{bias}}{n^2})$, reminding us of the bias-variance decomposition. In fact, it is possible to write such a decomposition for any Bregman divergence
Let $\tilde \theta_n := \expecta{\hat \theta_n}$ be the expectation of the MAP in primal space, and $\tilde \m_n = \nabla \logpart(\tilde \theta_n )$ be the corresponding mean parameter.
As described by \citet[Theorem 0.1]{pfau2013generalized}, the  expected Bregman decomposes like
\begin{align}
	\expect{\bregmanconj(\m^* ; \hat \m_n)} 
	&= {\bregmanconj(\m^* ; \tilde \m_n)}
	+ {\expect{\bregmanconj(\tilde \m_n ; \MAPm)}}
	\label{eq:primal_pivot}
\end{align}

@@

Relative smoothness has renewed interest for analyses of mirror descent
without relying on strong-convexity in an underlying norm.
Recent works have made attempts to generalize the assumptions 
that leads to the convergence of mirror descent 
with strongly-convex reference functions, 
such as a uniform bound on the variance of the stochastic gradients $\tilde g(\theta)$,
$\expecta{\norm{\nabla f(\theta) - \tilde g(\theta)}^2} \leq \sigma^2$, 
or only requiring this bounded variance at the minimum $\theta_*$.
However, they do not yet address the problem satisfactorily.
We present those generalization here 
and they are summarized in the context of maximum likelihood estimation 
in \cref{tbl:assumptions}.

@@

{\bf For a maximum likelihood problem, }
where $\tilde g(\theta) = \m - x$ 
where $x \sim p(x\cond\theta_*)$,
and $\nabla f(\theta) = \m - \m_*$,
\aligns{
	&
	\mathbb{E}_{x}\bigg[
	\cB_{A^*}\paren{(1-\eta) \m + \eta x, (1-\eta)\m + \eta \m_*}
	\\
	&\hphantom{\mathbb{E}_{\tilde g_t}[}
	\cB_{A^*}\paren{(1-\eta)\m + \eta \m_*, (1-\eta) \m + \eta x}
	\bigg]
	\leq 
	\eta C.
}
\fdk{How do we show that this fails?}
